---
title: "Homework 6"
author: Nikhita Damaraju
output: github_document
---

This is my solution to Homework 6.

Library imports 

```{r settings, message = FALSE} 
library(tidyverse)
library(dplyr)
library(stringr)
library(ggplot2)
library(modelr)


theme_set(theme_bw() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

# setting seed value
set.seed(1)
```

## Problem 1

Reading data and adding victim age variable
ß
```{r message = FALSE}
homicide_df = 
  read_csv("./datasets/homicide-data.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    city_state != "Tulsa, AL") %>%   # removing tulsa AL
  select(city_state, resolution, victim_age, victim_race, victim_sex)
```

### Logistic regression output for Baltimore

```{r}
baltimore_df =
  homicide_df %>% 
  filter(city_state == "Baltimore, MD")
glm(resolution ~ victim_age + victim_race + victim_sex, 
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```

Extracting adjusted odds ration for each

```{r}
models_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI")) 
```

```{r}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y = "estimated OR")
```

We observe that the city Albuquerque has the highest OR value compared to other cities. Further, it also has the longest CI interval range. The length of the confidence interval appears to be increasing towards the right.

## Problem 2

**Loading and cleaning dataset**

```{r message = FALSE}
weight_df = 
  read_csv('./datasets/birthweight.csv') %>%
  mutate(
    babysex = as.factor(case_when(
      babysex == 1 ~ 'male',
      babysex == 2 ~ 'female'
    )),
    
    frace = as.factor(case_when(
      frace == 1 ~ 'White',
      frace == 2 ~ 'Black',
      frace == 3 ~ 'Asian',
      frace == 4 ~ 'Puerto Rican',
      frace == 8 ~ 'Other',
      frace == 9 ~ 'Unknown')),
    
    mrace = as.factor(case_when(
      mrace == 1 ~ 'White',
      mrace == 2 ~ 'Black',
      mrace == 3 ~ 'Asian',
      mrace == 4 ~ 'Puerto Rican',
      mrace == 8 ~ 'Other',
      mrace == 9 ~ 'Unknown')),
    
    malform = as.factor(case_when(
      malform == 0 ~ 'absent',
      malform == 1 ~ 'present'
    ))
  ) %>%
  drop_na()
```

The dataset consisting of weights of children has 20 variables out of which, 4 variables - `babysex`, `frace`, `mrace` and `malform` are categorical. There are no missing values in the dataset and it consists of `r nrow(weight_df)` rows.

### Linear model

**Identifying predictors of interest**

For categorical variables, based on previous studies, there is an indication that ethnicity can influence the weight of the baby. For numerical variables, we look at the correlation with the outcome to identify variables that are independent correlating with the outcome. We also omit variables `pnulbw` and `pnumsga` due to large number of values with 0.

```{r}
# Selecting numerical variables 
weight_df %>%
  select(-c(babysex, frace, mrace, malform, pnumlbw, pnumsga)) %>%
  cor() %>% round(digits = 2) %>% knitr::kable()
```

Based on the results above, we pick our predictors as - bhead, blength, gaweeks, wtgain, babysex and the categorical variable - mrace

**Linear model**

```{r}
main_model = lm(bwt ~ bhead + blength + gaweeks + wtgain + mrace, data = weight_df)
```

**Model residuals against fitted values**

```{r message = FALSE}
weight_df %>%
  modelr::add_residuals(main_model) %>% 
  modelr::add_predictions(main_model) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  labs(x = 'Fitted values', y = 'Residuals', title = 'Residuals vs Fitted values')


```

### Comparison of model with other models 

```{r}
model1 = lm(bwt ~ blength + gaweeks, data = weight_df)
model2 = lm(bwt ~ bhead*blength*babysex, data = weight_df)

cv_df =
  crossv_mc(weight_df, 10) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>%
  mutate(
    main_model  = map(train, ~lm(bwt ~ bhead + blength + gaweeks + wtgain + mrace, data = .x)),
    model1 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    model2 = map(train, ~lm(bwt ~ bhead*blength*babysex, data = weight_df))) %>%
  mutate(
    rmse_main = map2_dbl(main_model, test, ~rmse(model = .x, data = .y)),
    rmse_model1 = map2_dbl(model1, test, ~rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(model2, test, ~rmse(model = .x, data = .y)))
```

**Cross-validated predicted error vs model of choice**

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  labs(x = 'model', y = 'RMSE', title = 'Cross-validated predicted error')
```

Based on the plot above, it appears that our model of choice `main_model` has the least mean RMSE out when compared to the other two models.

## Problem 3

Reading data

```{r message = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) 

weather_df = 
  weather_df %>%
  select(tmin, tmax)  # Selecting tmin and tmax as the only variables of interest
```

**Creating dataframe to draw 5000 bootstrapped samples and creating measures of interest**

### Distribution of r-squared

```{r message = FALSE}
bootstrap_rsq = 
  weather_df %>%
  modelr::bootstrap(n = 100) %>%
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance),
    ) %>%
  select(-strap, -models) %>%
  unnest(results) %>%
  select(r.squared)

# Plot 
bootstrap_rsq %>%
  ggplot(aes(x = r.squared)) +
  geom_histogram() +
  labs(title = "distribution of r.squared")

# 95% CI

```


### Distribution of log term

```{r}
bootstrap_log = 
  weather_df %>%
  modelr::bootstrap(n = 100) %>%
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy),
    ) %>%
  select(-strap, -models) %>%
  unnest(results) 

bootstrap_log_tidy = 
  bootstrap_log %>%
  select(-statistic, -p.value, -std.error) %>%
  group_by(.id) %>%
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) 

colnames(bootstrap_log_tidy) = c("id","B0","B1")

bootstrap_log_tidy = 
  bootstrap_log_tidy %>%
  mutate(
    log_term = log(B0*B1)
  ) 

bootstrap_log_tidy %>%
  ggplot(aes(x = log_term)) +
  geom_histogram() +
  labs(title = "distribution of log_term")
  
```



